Regression Cube Analysis of Cohort Study Data

### 2015-03-29 draft.tex

%% Steenwijk: Integrated Visual Analysis for Heterogeneous Datasets in Cohort Studies
Steenwijk et al. \cite{Steenwijk} focus on \emph{hypothesis-free} exploration of cohort data.
%%
They employ a framework consisting of feature extraction and visualization to derive dependencies between image- and non-image features.
%%
They incorporate linked views using scatter plots, bar charts, parallel coordinates and time plots to display and select the data.
%%
%Medical image data can be incorporated to investigate potentially interesting individuals.
%%
%%We distinguish to this and other linked-view analysis based approaches by showing abstracted information derived from a preceding analyze-first step to highlight relationships rather than displaying the actual data using various information visualization techniques.
%%
Generalized Pairs plots (\texttt{GPLOMÃ‚Â´S}) extend the idea of scatter plot matrices by the pairwise depiction of heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
%%
The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter and are therefore suitable for \emph{small} epidemiological data sets.
%%
Dai et al. \cite{Dai2005} incorporate a \texttt{GPLOM}-like visualization using choropleth maps mapping spatial data, such as mortality rates together with scatter plots augmented with \emph{Pearsson's} $r$ values.
%%
A \emph{concept map} summarizes features related to a specified disease.
%%
Time-dependent epidemiological data is visualized by Chui et al. \cite{Chui2011} using multi-panel graphs highlighting risk factor differences with age and gender towards influenza- and salmonellosis-associated hospitalizations.
%%
%We also include confounder features (age, gender) into the analysis while focusing on non-time-dependent cohort study data sets.
%%
\paragraph{Statistical Analysis.}
%%
%Statistical analysis is an inherent part of Visual Analysis.
%%
%Here we present work showing a strong focus on statistical analysis in this context.
%% Quality Metrics in High-Dimensional Data Visualization - An Overview and Systematization
Bertini et al. \cite{Bertini} present an overview over quality metrics describing high dimensional data.
%%
Their research agenda acts as guideline for our design, comprising applications using domain experts, perceptual tuning towards human pattern recognition of important aspects and scalability between different data set sizes.
%%
%% TODO: Use this one? "Statistical resilience of random populations to random perturbations" - http://www.ncbi.nlm.nih.gov/pubmed/19256997 - wie eine Variable sich ÃƒÂ¤ndert, wenn man per Zufall die Populationen verÃƒÂ¤ndert, um zu schauen wie zuverlÃƒÂ¤ssig diese Statistik ist.
%% The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems
Ahmadi et al. \cite{Ahmadi} define the \emph{Sparse Regression Cube}, which partitions sparse high-dimensional into subspaces, which are then described by their most reliable linear regression model.
%%
%In their application, they try to find most fuel-efficient roads connecting two user-defined landmarks.
%%
They focus on presenting an algebraic for efficient regression-model calculation to find the best fit for a subspace.
%%
%The dimensions for the calculations are fixed, while we assess different features for each model calculation and can therefore not benefit from a hierarchical calculation approach.

%% Hierarchical Brushing of High-Dimensional Data Sets Using Quality Metrics
Albuquerque et al. \cite{Albuquerque} present an interactive exploration framework displaying quality metrics of high dimensional data sets with brushing facilities to create subsets.
%%
%These subsets are then analyzed in detail, following a drilling-down approach.
%%
They analyzing subsets using a drilling-down approach by incorporating scatter plot matrices (\texttt{SPLOMÂ«S}) of quality metrics. %describing the features.
%%
%We incorporate the idea of abstracting features for comparison.
%%
Turkay et al. \cite{Turkay} follow a similar approach by using both descriptive metrics for features as well as the features themselves and incorporate them into linked plots.
%%
The approach was applied to a cognitive aging studies, where new hypotheses could be derived.

%Hielscher et al. \cite{Hielscher} assess subject similarity towards hepatic steatosis (fatty liver disease) to identify common risk factors using [...]?
%\com{Uli, can you please complete this, I do not have the pdf available!}
%% Learning and inspecting classification rules from longitudinal epidemiological data to identify predictive features on hepatic steatosis
Niemann et al. \cite{Niemann2014} investigate risk factors of the hepatic steatosis using decision trees with interactive data mining tools.
%%
They extracted association rules that serve as basis for a proof-of-concept evaluation.
%%
In \cite{Niemann2015}, Niemann et al. improve on classification performance by generating features (called \emph{Evolution Features}) that describe latent temporal information across the study waves.
%%
We try to reproduce their results and investigate their findings further.
%%
%\com{GlaÃƒÂŸer, S., Niemann, U., Preim, B., and Spiliopoulou, M. (2013). Can we Distinguish Between Benign and Ma- lignant Breast Tumors in DCE-MRI}

>>>>

%% Steenwijk: Integrated Visual Analysis for Heterogeneous Datasets in Cohort Studies
Steenwijk et al. \cite{Steenwijk} focus on \emph{hypothesis-free} exploration of cohort data.
%%
They employ a framework consisting of feature extraction and visualization to derive dependencies between image- and non-image features.
%%
They incorporate linked views using scatter plots, bar charts, parallel coordinates and time plots to display and select the data.
%%
%Medical image data can be incorporated to investigate potentially interesting individuals.
%%
%%We distinguish to this and other linked-view analysis based approaches by showing abstracted information derived from a preceding analyze-first step to highlight relationships rather than displaying the actual data using various information visualization techniques.
%%
Generalized Pairs plots (\texttt{GPLOMÃ‚Â´S}) extend the idea of scatter plot matrices by the pairwise depiction of heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
%%
The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter and are therefore suitable for \emph{small} epidemiological data sets.
%%
Dai et al. \cite{Dai2005} incorporate a \texttt{GPLOM}-like visualization using choropleth maps mapping spatial data, such as mortality rates together with scatter plots augmented with \emph{Pearsson's} $r$ values.
%%
A \emph{concept map} summarizes features related to a specified disease.
%%
Time-dependent epidemiological data is visualized by Chui et al. \cite{Chui2011} using multi-panel graphs highlighting risk factor differences with age and gender towards influenza- and salmonellosis-associated hospitalizations.
%%
%We also include confounder features (age, gender) into the analysis while focusing on non-time-dependent cohort study data sets.
%%
\paragraph{Statistical Analysis.}
%%
%Statistical analysis is an inherent part of Visual Analysis.
%%
%Here we present work showing a strong focus on statistical analysis in this context.
%% Quality Metrics in High-Dimensional Data Visualization - An Overview and Systematization
Bertini et al. \cite{Bertini} present an overview over quality metrics describing high dimensional data.
%%
Their research agenda acts as guideline for our design, comprising perceptual tuning towards human pattern recognition of important aspects, scalability between different data set sizes and application evaluation with domain experts.
%%
%% TODO: Use this one? "Statistical resilience of random populations to random perturbations" - http://www.ncbi.nlm.nih.gov/pubmed/19256997 - wie eine Variable sich ÃƒÂ¤ndert, wenn man per Zufall die Populationen verÃƒÂ¤ndert, um zu schauen wie zuverlÃƒÂ¤ssig diese Statistik ist.
%% The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems
Ahmadi et al. \cite{Ahmadi} define the \emph{Sparse Regression Cube}, which partitions sparse high-dimensional data into subspaces, which are then described by their most reliable linear regression model.
%%
%In their application, they try to find most fuel-efficient roads connecting two user-defined landmarks.
%%
They focus on an algebraic representation for efficient regression-model calculation to find the best fit for a subspace.
%%
%The dimensions for the calculations are fixed, while we assess different features for each model calculation and can therefore not benefit from a hierarchical calculation approach.

%% Hierarchical Brushing of High-Dimensional Data Sets Using Quality Metrics
Albuquerque et al. \cite{Albuquerque} present an interactive exploration framework displaying quality metrics of high dimensional data sets with brushing facilities to create subsets.
%%
%These subsets are then analyzed in detail, following a drilling-down approach.
%%
They analyze subsets using a drilling-down approach by incorporating scatter plot matrices (\texttt{SPLOMÂ«S}) of quality metrics. %describing the features.
%%
%We incorporate the idea of abstracting features for comparison.
%%
Turkay et al. \cite{Turkay} follow a similar approach by using both descriptive metrics for features as well as the features themselves and incorporate them into linked plots.
%%
The approach was applied to a cognitive aging studies, where new hypotheses could be derived.

%Hielscher et al. \cite{Hielscher} assess subject similarity towards hepatic steatosis (fatty liver disease) to identify common risk factors using [...]?
%\com{Uli, can you please complete this, I do not have the pdf available!}
%% Learning and inspecting classification rules from longitudinal epidemiological data to identify predictive features on hepatic steatosis
Niemann et al. \cite{Niemann2014} investigate risk factors of hepatic steatosis (fatty liver disease) using decision trees with interactive data mining tools.
%%
They extracted association rules that serve as basis for a proof-of-concept evaluation.
%%
In \cite{Niemann2015}, Niemann et al. improve on classification performance by generating features (called \emph{evolution features}), that describe latent temporal information across the study waves.
%%
We try to reproduce their results and investigate their findings further.
%%
%\com{GlaÃƒÂŸer, S., Niemann, U., Preim, B., and Spiliopoulou, M. (2013). Can we Distinguish Between Benign and Ma- lignant Breast Tumors in DCE-MRI}
